### 2015-07-06 narrative/data-description.tex

Our research will address the following questions:
1) How can users interact with the storage system efficiently?,
2) How can users express their intentions into actionable items which can take place in the middleware and/or storage layer?
3) How can users annotate data with a utility metric?

\paragraph{State of the art:}
Our current data annotation techniques demonstrated in
ADIOS~\cite{lofstead:2009:adaptible} implements a binary-packed data format that
allows data characteristics such as min, max and index to be wrapped around
data chunks. A direct benefit is that each data chunk can be operated upon
independently and I/O concurrently can be maximized. We will build upon this
capability and further augment the format to express data utility.  Recently
DAMSEL~\cite{damsel} provides a rich metadata representation and management
layer that captures the relationship between data blocks for scientific
applications.  This allows application data to be mapped to storage system
efficiently, without overburdening users to managing complex data model, such
as AMR, from the user space. However, the aspect of data importance hasn't been
explored, which we believe will be important for exascale storage solutions.
{\color{red}\bf there must be other work in this area, and we need to get more
references}

\paragraph{Proposed research approach:} 
We will design and develop new techniques for data management, description,
and access that allows users to describe the data utility based on their
expectations. This can be done by plugging user level functions that
partition user level data into chunks and be executed by the middleware to
calculate the utility value for each chunk. The utility value along with other
metadata will be encapsulated by the middleware and utilized by the underlying
storage system to management data placement.
%
%We will design an updated I/O API that will expand on the current POSIX I/O
%semantics by adding new information to the I/O request.

We will explore the set of application level hints that can be provided easily
to the storage system. These hints will enable the storage system to make the
most appropriate optimization decisions for this user in the context of the
multi-user environment rather than trying to rely on a one-size-fits-all
approach. In particular, we expect the application to provide hints on the
length of time before the output data is persistent and visible to users and
the expected data lifetime for output sets.  For read calls we expect the set
of hints to combine latency and precision requirements.  Additional hints will
also be investigated.

>>>>

Our research will address the following questions:
1) How can users interact with the storage system efficiently?
2) How can users express their intentions into actionable items which can take place in the middleware and/or storage layer?
3) How can users annotate data with a utility metric?

\paragraph{State of the art:}
Our current data annotation techniques demonstrated in
ADIOS~\cite{lofstead:2009:adaptible} implement a binary-packed data format that
allows data characteristics such as min, max and index to be wrapped around
data chunks. A direct benefit is that each data chunk can be operated upon
independently and I/O concurrency can be maximized. We will build upon this
capability and further augment the format to express data utility.
Recently
DAMSEL~\cite{damsel} has provided a rich metadata representation and management
layer that captures the relationship between data blocks for scientific
applications.  This allows application data to be mapped to storage system
efficiently, without overburdening users with the management of complex data models, such
as AMR, from the user space. Despite these efforts, the aspect of data importance has yet to be 
explored, and we believe it will be vital for exascale storage solutions.
{\color{red}\bf there must be other work in this area, and we need to get more
references}

\paragraph{Proposed research approach:} 
We will design and develop new techniques for data management, description,
and access that allow users to describe the data utility based on their
expectations. This can be done by allowing users to plug in functions that
partition user level data into chunks, which are then executed by the middleware to
calculate the utility value for each chunk. This value, along with other
metadata, will be encapsulated by the middleware and utilized by the underlying
storage system to manage data placement.
%
%We will design an updated I/O API that will expand on the current POSIX I/O
%semantics by adding new information to the I/O request.

We will explore the space of application level hints that can be provided easily
to the storage system. These hints will enable the storage system to make the
most appropriate optimization decisions for this user in the context of the
multi-user environment rather than trying to rely on a one-size-fits-all
approach. In particular, we expect the application to provide hints on the
length of time before the output data is persistent and visible to users and
the expected data lifetime for output sets. For read calls we envision
hints that provide latency and precision requirements. Additional hints will
also be investigated.
